\chapter*{Conclusion}
\chaptermark{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}

The exploration of digital art collections using Link-Traversal–based Query Processing has been a multifaceted journey, intertwining the realms of art and technology. As digital art collections have become more accessible, challenges have arisen, particularly for individuals without a technical background. This research embarked on addressing these challenges, aiming to provide tools and methodologies that empower both professionals and art enthusiasts to delve deeper into the digital art landscape, making new discoveries and drawing meaningful connections.

The systematic process of discovering digital art collections can be broadly categorized into three main steps: building queries, executing them — with a specific focus on link traversal in this research — and handling the results through visualization and storage. Each chapter of this research has meticulously addressed one of these steps.

Chapter~\ref{chap:coghent_link_traversal} delved into the execution of queries using Link-Traversal-based Query Processing. The chapter emphasized the potential of link traversal in uncovering specific attributes of the Collections of Ghent's Human-Made Objects, offering insights that would remain hidden with traditional querying methods. However, it also highlighted the inherent challenges and unpredictability associated with link traversal. For instance, due to server misconfigurations, certain resources like Stad Gent's could not be accessed, while the use of others', like Getty Vocabularies', required workarounds. These challenges underscored the fragility of Link Traversal-based Query Processing compared to traditional SPARQL querying.

Chapter~\ref{chap:tools_query_building} introduced tools that simplify the intricate task of query formulation, making it more accessible to a broader audience. While these tools are valuable, they are not presented as the definitive solutions. Instead, their core query-building functionality is designed to be modular, allowing others to adapt and use it for their own discovery applications.

Chapter~\ref{chap:handling_query_results} addressed the post-query phase, focusing on the visualization and preservation of query results. The chapter weighed the advantages and disadvantages of different visualization and preservation methods. However, it also highlighted areas for further exploration, such as the adaptability of visualization tools to real-time query updates and the potential for presenting query results as immersive, interactive narratives. These areas present exciting avenues for future research.

In conclusion, this research has provided valuable insights and tools for the discovery process of digital art collections, while also shedding light on the inherent challenges of the process. Specifically, the fragility of link traversal, its slower performance compared to traditional SPARQL querying, and the unpredictability of outcomes due to server misconfigurations or other unforeseen technical issues, pose significant hurdles. However, the undeniable potential of link traversal to uncover hidden data and offer deeper insights into digital art collections offers a promising future. As technology evolves and these challenges are addressed, it is anticipated that link traversal and its associated tools will become more mainstream, benefiting a wider audience. Yet, it is important to note that, at the time of this research's publication, harnessing its full capabilities still demands a certain level of technical expertise.

\phantomsection
\section*{Ethical and social reflection}
\addcontentsline{toc}{section}{Ethical and social reflection}

In examining the ethical and societal dimensions of the research, it is instructive to consider its alignment with the 17 \textit{Sustainable Development Goals}\footnote{\url{https://sdgs.un.org/goals}} (SDG) of the United Nations' \textit{2030 Agenda for Sustainable Development}\footnote{\url{https://sdgs.un.org/2030agenda}}.

Primarily, the research's emphasis on open-source applications seeks to broaden access to the discovery of art and culture, irrespective of technical expertise. This approach has parallels with SDG 4\footnote{\url{https://sdgs.un.org/goals/goal4}}, which highlights inclusive education, and SDG 10\footnote{\url{https://sdgs.un.org/goals/goal10}}, which addresses the reduction of inequalities.

Furthermore, the use of Linked Open Data in the research provides a mechanism for global cultural representation. With the study primarily focusing on Ghent's cultural heritage, its objectives intersect with SDG 11\footnote{\url{https://sdgs.un.org/goals/goal11}}, which is concerned with the preservation of global cultural heritage. The collaborative nature of the CoGhent partnership also reflects the spirit of SDG 17\footnote{\url{https://sdgs.un.org/goals/goal17}}, which underscores the importance of partnerships.

The research's utilization of the semantic web offers a method for accessible art collection discovery, which can be related to SDGs 8\footnote{\url{https://sdgs.un.org/goals/goal8}} and 9\footnote{\url{https://sdgs.un.org/goals/goal9}}, centered around economic growth and innovation. Additionally, the semantic web's approach to data control aligns with SDG 16's\footnote{\url{https://sdgs.un.org/goals/goal16}} emphasis on transparency and accountability.

In summary, although the research does not \textit{directly} address \textit{all} the SDGs mentioned above, its utilization of the semantic web and the introduction of various open-source tools at least provide \textit{avenues} for potentially aligning with them.